---
title: "Simulation (reference) model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sim-model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(impactflu)
library(DiagrammeR)
```

## Compartments

```{r, echo = FALSE}
model <- "sim-model.viz"
grViz(model, width = 500, height = 500)
```

$d$ --- day index.\
$g$ --- lag period.\
$w_d$ --- vaccine effectiveness on day $d$.\
$v_d$ --- vaccine coverage on day $d$.\
$r_d$ --- flu incidence on day $d$.\
$m_d$ --- flu mortality on day $d$.\
$b_d$ --- number of vaccinations of susceptible on day $d$.\
$e_d$ --- number of infections of non-vaccinated on day $d$.\
$f_d$ --- number of infections of vaccinated on day $d$.\

## Calculations

$d=0$ --- initial conditions.\
$d=1$ --- first timepoint (day) for which we have data.\
$N$ --- starting population size (given).\
$V_d$ --- vaccinations on day $d$ (given).\
$S_d^*$ --- cases on day $d$ in absence of vaccination (given).\
$e_d$ --- vaccine effectiveness on day $d$ (given).\
$P_d^*$ --- non-cases on day $d$ in absence of vaccination.\
$S_d^a$ --- averted cases on day $d$.\
$S_d$ --- cases on day $d$.\
$Q_d^*$ --- deaths on day $d$ in absence of vaccination (given).\
$Q_d$ --- deaths on day $d$.\
$Q_d^a$ --- averted deaths on day $d$.\

For $d=0$:

$A_0 = N$\
$B_0 = 0$\
$C_0 = 0$\
$D_0 = 0$\
$E_0 = 0$\
$F_0 = 0$\
$G_0 = 0$\
$H_0 = 0$\
$P_0^* = N$\
$S_0 = 0$\
$S_0^a = 0$\

For $d>0$ (order presented corresponds to the order of calculations):

$r_d = \frac{S_d^*}{P_{d-1}^*}$\
$m_d = \frac{Q_d^*}{P_{d-1}^*}$\
$S_d=r_dA_{d-1} + r_dC_{d-1} + r_dB_{d-1}$\
$Q_d=m_dE_{d-1} + m_dF_{d-1}$\
$P_d^* = P_{d-1}^* - S_d^*$\
$S_d^a = S_d^* - S_d$\
$v_d = \frac{V_d}{A_{d-1}+E_{d-1}}$\
$b_d=v_dA_{d-1}$\
$A_d = A_{d-1} - r_dA_{d-1}-b_d$\
$b_{d-1} ... b_{d-g}=b_{d-1} ... b_{d-g} - r_db_{d-1} ... b_{d-g}$\
$B_d = B_{d-1} - r_dB_{d-1} + b_d - b_{d-g}$\
$C_d = C_{d-1} - r_dC_d + b_{d-g} - e_db_{d-g}$\
$D_d = D_{d-1} + w_db_{d-g}$\
$E_d = E_{d-1} + w_dA_{d-1} - v_dE_{d-1}$\
$F_d = F_{d-1} + r_dC_{d-1} + r_dB_{d-1} + v_dE_{d-1}$\
$G_d = G_{d-1} + m_dE_{d-1}$\
$H_d = H_{d-1} + m_dF$\

Line $b_{d-1} ... b_{d-g}$ means that the same calculation is done on each $b_{d-i}$ from $i=1$ to $i=g$. The calculation being $b_{d-i}=b_{d-i}-r_db_{d-i}$.

When simualtions are random, each multiplication of a compartment by a probability e.g. $r_dA_{d-1}$, becomes a binomial random number e.g. $\text{Binomial}(A_{d-1}, r_d)$).
